Repeat Yourself? When Linux kernel maintainers balance Code Duplication and Readability

### 2025-10-20 paper/techdebt2026/techdebt.tex
techdebt2025: Updates abstract
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The vast scale and continuous evolution of the Linux kernel make its maintenance a complex undertaking, where code duplication remains a persistent challenge that can hinder development and introduce bugs. This paper presents a multimethod ethnographic study investigating the socio-technical dynamics of contributing to the Linux kernel by mitigating identified redundancies. To support this investigation, we developed \texttt{anonymous-tool}, a command-line utility for detecting function-level duplications, offering a concrete entry point for new contributors. Our ethnographic approach, including participant and non-participant observations, demonstrated that addressing duplications is viable for lowering the contribution barrier. This point is evidenced by the main author and 7 of 11 (64\%) student groups having their patches accepted into the kernel, collectively removing 991 lines of duplicated code. Moreover, the study also reveals a more complex reality beyond simple clone removal: an analysis of maintainer feedback on both accepted and rejected patches highlights a nuanced understanding of code quality within the Linux kernel community, where the benefits of eliminating duplication are carefully weighed against factors such as readability, the introduction of new abstractions, and the specific context of the code.
\end{abstract}

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The vast scale and continuous evolution of the Linux kernel make its maintenance a complex undertaking, where code duplication remains a persistent challenge that can hinder development and introduce bugs. This paper presents a multimethod ethnographic study investigating the socio-technical dynamics of contributing to the Linux kernel by mitigating identified redundancies. To support this investigation, we developed \texttt{anonymous-tool}, a command-line utility for detecting function-level duplications, offering a concrete entry point for new contributors. Our ethnographic approach, including participant and non-participant observations, demonstrated that addressing duplications is viable for lowering the contribution barrier. This point is evidenced by the main author and 7 of 11 (64\%) student groups having accepted contributions in the Linux project, collectively removing 991 lines of duplicated code. Nevertheless, the study reveals a more intriguing and complex reality beyond mindlessly removing clones. An analysis of maintainer feedback on both accepted and rejected contributions highlights a nuanced understanding of technical debt due to code duplication within the Linux kernel community, where the benefits of eliminating duplications are carefully weighed against factors such as readability, the introduction of new abstractions, and the specific code context.
\end{abstract}
